\chapter{System Implementation}
\label{chap:ch3}

\tolerance=9999
\emergencystretch=3em

\par This chapter describes how the peer-assisted streaming system was built in
practice.  It begins with the server-side components, covering the data model,
the streaming endpoints, the peer registry, and the signaling server.  It then
moves to the client-side implementation in Flutter, where the focus is on how
the audio playback library was chosen, what limitations were discovered on each
platform, how the platform differences were bridged, and how the chunk retrieval
and WebRTC subsystems work in detail.  The chapter ends with annotated flow
diagrams that show how the individual components fit together at runtime.

% ─────────────────────────────────────────────────────────────────────────────
\section{Backend Data Model and Storage Layer}
\label{sec:data-model}

\par The persistence layer is built on PostgreSQL and managed through Liquibase
changesets, keeping the schema under version control and decoupled from the JPA
entity definitions.  The JPA DDL mode is set to \texttt{validate-only}, so every
structural change must be expressed as a new changeset rather than relying on
automatic schema generation.  This avoids the common problem of a running
application silently dropping or altering columns during startup.

\subsection{Content-Addressed Chunk Storage}
\label{subsec:chunk-storage}

\par The central design decision for audio storage is to decompose each song into
fixed-size \emph{chunks} of exactly 65\,536 bytes (64~KiB) and to store them
independently.  Each chunk is represented by the \texttt{Chunk} entity, whose most
important field is \texttt{contentHash}, a 256-bit SHA-256 digest of the raw audio
bytes.  The hash serves a dual purpose: it is a natural deduplication key, so that
two songs sharing an identical segment only ever store one physical file, and it
provides an integrity fingerprint that both the server and the client can verify
before using the data.

\par The \texttt{storagePath} field records the filesystem location of the
corresponding file, and the \texttt{size} field caches the byte count for
range-request arithmetic.  The \texttt{SongChunk} join entity binds a \texttt{Song}
to its constituent \texttt{Chunk} objects via an ordered \texttt{orderIndex}
column, preserving the sequential structure of the audio file.  The relationship is
illustrated in Figure~\ref{fig:chunk-model}.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    entity/.style={rectangle, draw, rounded corners=3pt,
                   minimum width=3.2cm, minimum height=0.9cm,
                   font=\small\ttfamily},
    field/.style={rectangle, draw=gray!50, minimum width=3.2cm,
                  minimum height=0.7cm, font=\footnotesize\ttfamily,
                  fill=gray!8},
    lbl/.style={font=\scriptsize, midway, above}
]
  % Song
  \node[entity, fill=blue!10] (song) at (0,0) {Song};
  \node[field, below=0pt of song, yshift=-0.05cm] (s1) {id : Long};
  \node[field, below=0pt of s1] (s2) {fileHash : String};
  \node[field, below=0pt of s2] (s3) {durationInSeconds : int};
  \node[field, below=0pt of s3] (s4) {ownerId : Long?};

  % SongChunk
  \node[entity, fill=orange!15] (sc) at (5.5,0) {SongChunk};
  \node[field, below=0pt of sc, yshift=-0.05cm] (c1) {id : Long};
  \node[field, below=0pt of c1] (c2) {orderIndex : Integer};

  % Chunk
  \node[entity, fill=green!10] (ch) at (11,0) {Chunk};
  \node[field, below=0pt of ch, yshift=-0.05cm] (ch1) {id : Long};
  \node[field, below=0pt of ch1] (ch2) {contentHash : String};
  \node[field, below=0pt of ch2] (ch3) {storagePath : String};
  \node[field, below=0pt of ch3] (ch4) {size : Integer};

  % Arrows
  \draw[-{Stealth}] (s4.east) -- node[above, font=\scriptsize]{1..*} (c1.west);
  \draw[-{Stealth}] (c2.east) -- node[above, font=\scriptsize]{*..1} (ch1.west);
\end{tikzpicture}
\caption{Simplified entity relationship between \texttt{Song}, \texttt{SongChunk}, and \texttt{Chunk}.}
\label{fig:chunk-model}
\end{figure}

\par Because chunks are content-addressed, the upload pipeline hashes each
64~KiB block before writing it to disk.  If an identical hash already exists in
the database, the physical file is reused and only a new \texttt{SongChunk} row is
inserted.  This strategy is particularly effective for music libraries where
intros, outros, and silence segments repeat across many tracks.

\subsection{User Library and Soft Deletion}
\label{subsec:user-library}

\par Each user's relationship to a song is captured by the \texttt{UserLibrary}
entity, which uses a composite primary key of \texttt{(userId, songId)}.
Engagement fields, namely \texttt{liked}, \texttt{playCount}, \texttt{lastPlayed},
and \texttt{addedAt}, are stored alongside a \texttt{lastUpdated} timestamp that
drives the incremental synchronisation protocol described in
Section~\ref{sec:data-sync}.  Deletion is implemented as a soft flag
(\texttt{isDeleted}) so that the sync engine can propagate removal events to other
devices without losing the record itself.

\par The \texttt{ownerId} field on \texttt{Song} distinguishes between
platform-owned catalogue tracks (nullable owner) and user-uploaded files (set
owner).  The streaming service enforces this at the query level: requests for
user-uploaded songs are rejected if the authenticated user does not match
\texttt{ownerId}, providing simple but effective access control.

% ─────────────────────────────────────────────────────────────────────────────
\section{Server-Side Streaming Service}
\label{sec:streaming-service}

\par The \texttt{StreamingService} exposes three retrieval modes to REST clients:
a manifest endpoint, a per-chunk endpoint, and a prefix endpoint.  All three are
generated from the OpenAPI specification in \texttt{api.yml}.  The service class
implements the generated controller interface and contains the business logic,
keeping the controllers thin.

\subsection{Chunk Manifest}
\label{subsec:manifest}

\par Before a client can stream a song it fetches the \emph{chunk manifest}: a
lightweight JSON document that describes the song's chunk count, the size of each
individual chunk, the total byte length, and an ordered list of SHA-256 hashes,
one per chunk.  The manifest allows the client to verify every chunk it receives
regardless of whether that chunk arrived from the origin server or from a peer.
Generating the manifest requires a single database join (\texttt{Song} $\rightarrow$
\texttt{SongChunk} $\rightarrow$ \texttt{Chunk}) ordered by \texttt{orderIndex},
after which the service projects the \texttt{contentHash} fields into the response
DTO.

\subsection{Prefix Delivery}
\label{subsec:prefix}

\par To guarantee that playback can begin immediately, before any peer connection
is established, the server supports a \emph{prefix endpoint} that returns the first
$N$ bytes of a song as a single binary response.  The implementation iterates
through the ordered \texttt{SongChunk} list, opens each chunk file, appends bytes
to a buffer, and returns once the threshold is reached.  The exact number of prefix
chunks is computed dynamically on the client as described in
Section~\ref{subsec:prefix-calc}, but the server simply returns as many bytes as
requested, making it easy to adjust the threshold without a server deployment.

\subsection{Per-Chunk and Full-Stream Endpoints}
\label{subsec:per-chunk}

\par Individual chunk retrieval (\texttt{getSongChunk}) looks up the
\texttt{SongChunk} for the requested \texttt{(songId, chunkIndex)} pair, opens the
physical file at \texttt{storagePath}, and returns the raw bytes.  The full-stream
endpoint assembles a \texttt{SequenceInputStream} over all chunk files and is
provided as a compatibility fallback for clients that cannot support the chunked
protocol, for example certain browser-embedded players.

% ─────────────────────────────────────────────────────────────────────────────
\section{Peer Registry with Redis}
\label{sec:peer-registry}

\par The server must know, at the moment a client requests peer discovery, which
other clients hold which chunks.  This state is ephemeral, it exists only while
clients are connected, and must be shared across multiple server instances in a
horizontally-scaled deployment.  Redis was chosen for this purpose because it
provides the required atomicity primitives (hash operations, set operations) with
sub-millisecond latency.

\par The \texttt{PeerTrackingService} maintains two data structures in Redis for
each active peer:

\begin{itemize}
  \item \textbf{Chunk index}: a Redis \emph{hash} keyed as
        \texttt{peer:chunks:\{songId\}}, mapping each peer identifier to a
        JSON-serialised \texttt{Set<Integer>} of chunk indices that the peer has
        locally cached.
  \item \textbf{Song index}: a Redis \emph{set} keyed as
        \texttt{peer:songs:\{peerId\}}, recording which song IDs the peer has
        registered chunks for.  This reverse index enables efficient cleanup on
        disconnection.
\end{itemize}

\par When a peer announces a new set of cached chunks (message type
\texttt{REGISTER\_CACHE}), the service fetches the existing entry for that
\texttt{(songId, peerId)} pair, computes the union with the incoming indices, and
writes the result back.  This incremental merge avoids requiring peers to
retransmit their full cache state on every update.

\par On disconnection, \texttt{unregisterPeer} iterates the song index to enumerate
every song the peer had cached, removes the peer's entry from each chunk hash, and
then deletes the song index itself, leaving no stale entries in Redis.

% ─────────────────────────────────────────────────────────────────────────────
\section{WebSocket Signaling Server}
\label{sec:signaling}

\par WebRTC connections between clients require out-of-band signaling to exchange
Session Description Protocol (SDP) offers and answers as well as Interactive
Connectivity Establishment (ICE) candidates.  The \texttt{SignalingHandler}
implements the Spring \texttt{TextWebSocketHandler} interface and acts as this
signaling relay, while also serving as the delivery channel for server-initiated
events such as synchronisation triggers and playback state changes.

\subsection{Connection State}
\label{subsec:conn-state}

\par The handler maintains three in-memory concurrent maps:

\begin{itemize}
  \item \texttt{registry}: maps each WebSocket session ID to a
        \texttt{ClientConnection} record, which bundles the
        \texttt{WebSocketSession}, the peer identifier, and the optional
        authenticated user ID.
  \item \texttt{userIndex}: maps each user ID to the set of session IDs belonging
        to that user, allowing server events to fan out to all of a user's connected
        devices.
  \item \texttt{peerIndex}: maps each peer identifier to its session ID, enabling
        direct routing when the target peer is hosted by this server instance.
\end{itemize}

\par When a new connection arrives, a preliminary \texttt{ClientConnection} is
created with placeholder identifiers.  The first incoming message that carries a
\texttt{senderId} and \texttt{userId} promotes the connection to a fully identified
state, populating both indexes.

\subsection{Message Routing}
\label{subsec:msg-routing}

\par The message types the handler processes are listed below:

\begin{itemize}
  \item \textbf{\texttt{REGISTER\_CACHE}}: Delivered by a peer after it has cached
        one or more chunks.  The payload is a set of integer chunk indices.  The
        handler delegates to \texttt{PeerTrackingService.registerPeerChunks}.

  \item \textbf{\texttt{DISCOVER\_PEERS}}: A request for the chunk maps of all other
        peers that hold data for a given song.  The handler calls
        \texttt{PeerTrackingService.getPeerBufferMapsForSong}, filters out the
        requesting peer's own entry, and replies with a \texttt{PEER\_BUFFER\_MAP}
        message.

  \item \textbf{\texttt{OFFER} / \texttt{ANSWER} / \texttt{ICE\_CANDIDATE}}:
        WebRTC signaling messages that carry a \texttt{targetId} field.  The handler
        attempts to route the message to the target peer locally via
        \texttt{peerIndex}; if the target session is not present on this instance,
        the message is published to the Redis \texttt{signaling:webrtc} channel so
        that a \texttt{RedisSignalingListener} on another instance can deliver it.
\end{itemize}

\par The Redis pub/sub bridge enables horizontal scaling: each server instance runs
a \texttt{RedisSignalingListener} that subscribes to three channels,
\texttt{signaling:webrtc}, \texttt{signaling:sync}, and
\texttt{signaling:playback}, and delivers arriving messages to the locally
connected sessions they target.  Figure~\ref{fig:signaling-flow} illustrates the
cross-instance routing path.

\begin{figure}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=1.4cm and 2.8cm,
  box/.style={rectangle, draw, rounded corners=4pt,
              minimum width=2.6cm, minimum height=0.85cm,
              font=\small, align=center},
  redis/.style={box, fill=red!15},
  srv/.style={box, fill=blue!12},
  cli/.style={box, fill=green!12},
  arr/.style={-{Stealth}, thick}
]
  \node[cli]  (peerA) {Peer A\\(Client)};
  \node[srv, right=of peerA] (srv1) {Server\\Instance 1};
  \node[redis, right=of srv1] (redis) {Redis\\Pub/Sub};
  \node[srv, right=of redis] (srv2) {Server\\Instance 2};
  \node[cli, right=of srv2] (peerB) {Peer B\\(Client)};

  \draw[arr] (peerA) -- node[above, font=\scriptsize]{OFFER} (srv1);
  \draw[arr] (srv1)  -- node[above, font=\scriptsize]{publish} (redis);
  \draw[arr] (redis) -- node[above, font=\scriptsize]{subscribe} (srv2);
  \draw[arr] (srv2)  -- node[above, font=\scriptsize]{OFFER} (peerB);
\end{tikzpicture}%
}
\caption{Cross-instance WebRTC signal routing via Redis pub/sub.}
\label{fig:signaling-flow}
\end{figure}

% ─────────────────────────────────────────────────────────────────────────────
\section{Data Synchronisation Service}
\label{sec:data-sync}

\par The \texttt{DataSyncService} implements a bidirectional, timestamp-based
synchronisation protocol that keeps the user's library consistent across devices
without requiring a full reload of the catalogue.

\par A client sends a \texttt{SyncRequestDto} containing an optional
\texttt{lastSyncTime} ISO-8601 timestamp and a list of \texttt{localChanges}: each
one a \texttt{SongSyncDto} encoding a song's updated engagement fields
(\texttt{liked}, \texttt{playCountDelta}, \texttt{lastPlayed}, \texttt{isDeleted}).
The service applies these changes to the corresponding \texttt{UserLibrary} rows
inside a single transaction, then queries for all \texttt{UserLibrary} entries with
\texttt{lastUpdated > lastSyncTime} to collect server-side changes.  The response
carries both sets of changes and a new \texttt{syncTimestamp} that the client
stores locally for use in the next synchronisation call.

\par After the write phase completes, the service invokes
\texttt{SignalingHandler.sendSyncTrigger(userId)}, publishing a notification to the
Redis \texttt{signaling:sync} channel.  All other devices belonging to the same
user receive this trigger via the pub/sub bridge and initiate their own sync
request, propagating library changes across devices in near real time.

% ─────────────────────────────────────────────────────────────────────────────
\section{Frontend Framework and Dependency Injection}
\label{sec:frontend-di}

\par The client application is built with Flutter, a cross-platform UI framework
developed by Google that compiles to native code on desktop and mobile and to
JavaScript-based WebAssembly on the browser.  Flutter was chosen over alternatives
such as React Native or Electron primarily because it produces a single codebase
that targets Linux, macOS, Windows, Android, iOS, and the web browser without
requiring separate rendering engines or bridging layers for each platform.  This
matters for this project because the P2P streaming system requires the same chunk
retrieval logic, the same signaling protocol, and the same integrity verification
to run on every platform.  Duplicating that logic in separate codebases would
introduce an unacceptable maintenance burden.

\par A secondary reason for choosing Flutter is that the \texttt{flutter\_webrtc}
package provides a consistent WebRTC API across all supported targets.  On native
platforms it wraps the platform's native WebRTC library; on the web it delegates to
the browser's built-in WebRTC implementation.  This means the signaling logic and
the data channel protocol described in Section~\ref{sec:webrtc} do not need to
change depending on the platform.

\subsection{Provider-Based Dependency Injection}
\label{subsec:di}

\par State management and dependency injection are handled by the \texttt{provider}
package, which implements an InheritedWidget-based service locator pattern.  All
major services, such as \texttt{AppAudioService}, \texttt{WebRTCService},
\texttt{ChunkService}, and the various REST clients, are registered at the root of
the widget tree inside the \texttt{AbstractApp} base class.  Child widgets and
services retrieve their dependencies by calling \texttt{context.read<T>()}, which
walks up the widget tree and returns the nearest registered instance.

\par The dependency graph is constructed once at startup and the services are
long-lived for the duration of the application session.  The only exception is
\texttt{ChunkService}, which is instantiated per-song because its state (manifest,
RAM cache, active request deduplication map, peer completers) is specific to a
single audio file.  \texttt{AppAudioService} holds a factory function
\texttt{createChunkManager(fileHash)} that creates a new \texttt{ChunkService}
instance on demand.  This factory is injected at startup so that the audio service
never has a compile-time dependency on any particular \texttt{ChunkService}
implementation, which also makes unit testing straightforward.

\subsection{Platform-Specific Repository Split}
\label{subsec:repo-split}

\par The application targets platforms with fundamentally different storage
capabilities.  Desktop and mobile operating systems expose a persistent filesystem
and support native database libraries such as ObjectBox.  Web browsers, however,
do not expose a filesystem to web applications and cannot run native binary
libraries.

\par To handle this difference, the application defines a set of abstract
repository interfaces (for example \texttt{SongRepository},
\texttt{ChunkCacheRepository}, \texttt{SettingsRepository}) and provides two
implementations for each:

\begin{itemize}
  \item \textbf{ObjectBox-backed} implementations used on Linux, macOS, Windows,
        Android, and iOS.  ObjectBox is a high-performance embedded NoSQL database
        that maps Dart objects directly to binary storage with no SQL layer.  It was
        chosen over SQLite because it requires no schema migrations at the
        application level, schema changes are handled by dropping and rebuilding the
        store, and because its Dart code generation tooling integrates cleanly with
        Flutter's build pipeline.  The chunk cache itself is stored as raw binary
        files on disk rather than inside ObjectBox, because storing large binary
        blobs in a database can cause memory pressure during reads; the
        \texttt{IOChunkCacheRepository} simply writes each chunk to a per-song
        subdirectory and reads it back by file path.

  \item \textbf{In-memory} implementations used on the web.  These store all data
        in \texttt{Map} and \texttt{List} structures that live for the duration of
        the browser session.  Because browser tab memory is limited and there is no
        persistent storage permission needed, the in-memory approach is the simplest
        safe choice.  The trade-off is that the web platform cannot act as a strong
        P2P seeder between sessions because locally cached chunks are lost when the
        tab is closed.
\end{itemize}

\par The platform-specific app classes (\texttt{LinuxApp}, \texttt{WebApp}, and
their equivalents for other platforms) extend a shared \texttt{AbstractApp} base
class and override \texttt{platformProviders()}, injecting the correct repository
implementations before the common providers are registered.  This means the rest
of the application code never needs a platform check to decide which repository to
use; it always receives the correct one through the Provider tree.
Figure~\ref{fig:di-split} illustrates this separation.

\begin{figure}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=0.9cm and 2.2cm,
  base/.style={rectangle, draw, fill=blue!10, rounded corners=4pt,
               minimum width=5cm, minimum height=0.85cm,
               font=\small, align=center},
  plat/.style={rectangle, draw, fill=orange!15, rounded corners=4pt,
               minimum width=3.8cm, minimum height=0.75cm,
               font=\small, align=center},
  impl/.style={rectangle, draw, fill=green!12, rounded corners=4pt,
               minimum width=3.8cm, minimum height=0.75cm,
               font=\small, align=center},
  arr/.style={-{Stealth}, thick}
]
  \node[base] (abs) {AbstractApp\\(common providers, services)};

  \node[plat, below left=1.2cm and 3.5cm of abs]  (lin)
        {LinuxApp / MacApp\\WindowsApp / AndroidApp / IOSApp};
  \node[plat, below right=1.2cm and 3.5cm of abs] (web) {WebApp};

  \node[impl, below=0.7cm of lin] (objbox)
        {ObjectBox repos\\IOChunkCacheRepository};
  \node[impl, below=0.7cm of web] (mem)
        {InMemory repos\\InMemoryChunkCacheRepository};

  \draw[arr] (abs.south west) -- (lin.north);
  \draw[arr] (abs.south east) -- (web.north);
  \draw[arr] (lin.south) -- (objbox.north);
  \draw[arr] (web.south) -- (mem.north);
\end{tikzpicture}%
}
\caption{Platform repository split: ObjectBox for native targets, in-memory for web.}
\label{fig:di-split}
\end{figure}

% ─────────────────────────────────────────────────────────────────────────────
\section{Audio Playback Library: Choosing just\_audio}
\label{sec:audio-library}

\par The first decision to make when building a Flutter music player is which
audio playback library to use.  The most popular option in the Flutter ecosystem
is \texttt{just\_audio}, a package developed by Ryan Heise that wraps the native
audio APIs of each platform behind a unified Dart interface.  \texttt{just\_audio}
was chosen as the starting point for several reasons:

\begin{itemize}
  \item It supports a custom \texttt{StreamAudioSource} abstract class that lets
        the application provide audio bytes on demand, byte range by byte range,
        rather than passing a pre-downloaded file or a static URL.  This interface
        is exactly what the chunk-by-chunk retrieval pipeline needs.

  \item It exposes a \texttt{processingStateStream} and a \texttt{positionStream}
        that the application can observe to detect stalls and implement the
        watchdog timer described in Section~\ref{subsec:stuck}.

  \item It handles gapless playback, seeking, speed control, and pitch shifting
        through a single API surface that is consistent across Android, iOS, and
        the web.

  \item It integrates with the \texttt{audio\_service} package, which exposes
        media controls to the operating system (for example, the lock screen media
        controls on Android and iOS, and the MPRIS interface on Linux).
\end{itemize}

\par The key abstraction that makes the P2P streaming work with \texttt{just\_audio}
is the \texttt{StreamAudioSource} interface.  When the audio player needs data, it
calls the \texttt{request(start, end)} method on the source object, passing the
byte range it wants.  The source is responsible for returning a
\texttt{StreamAudioResponse} that contains the content length and a
\texttt{Stream<List<int>>} that yields the requested bytes.  The application
implements this interface in \texttt{P2PChunkedAudioSource}, which translates each
byte-range request into one or more calls to \texttt{ChunkService.getChunk}, slices
the chunk bytes to match the exact range, and yields them to the stream.  This
translation is described in detail in Section~\ref{sec:audio-source}.

\subsection{Limitations on Desktop Platforms}
\label{subsec:just-audio-desktop}

\par During development it became clear that \texttt{just\_audio} has significant
limitations on Linux and Windows.  The library relies on GStreamer on Linux and on
the Windows Media Foundation on Windows for its low-level audio decoding.  Both of
these backends have limited support for the custom \texttt{StreamAudioSource}
interface: they expect a seekable data source, such as a file path or a URL, and
cannot always handle the dynamic, on-demand nature of the chunk stream.

\par The most common problem encountered was that seeking, for example when the
user drags the progress bar, caused the player to call \texttt{request} with a
range that the underlying backend could not service because it expected the previous
range to be fully delivered first.  This produced audio glitches and, in some
cases, caused the player to stall permanently and require a full source reload.
The issue is not a bug in \texttt{just\_audio} itself but rather a limitation of
the GStreamer and Windows Media Foundation pipeline models, which are built around
the assumption that data arrives in a sequential stream rather than in arbitrary
ranges.

\par To work around this, the application switches to the \texttt{media\_kit}
package on Linux, macOS, and Windows.  \texttt{media\_kit} is a Dart/Flutter
wrapper around the \texttt{libmpv} media player library, which in turn uses FFmpeg
for decoding.  \texttt{libmpv} has a more flexible streaming model that can handle
non-linear seeking over a custom byte source because its internal demuxer is
designed around random-access I/O from the start.  The package exposes a player
API similar to \texttt{just\_audio} but tailored to desktop use cases.

\par The trade-off is that \texttt{media\_kit} does not implement the
\texttt{StreamAudioSource} abstraction that \texttt{just\_audio} defines, so the
chunk source had to be adapted for desktop.  On desktop platforms the audio source
is provided as a custom URI scheme that \texttt{libmpv} calls back into the Dart
application to read bytes from.  The rest of the chunk retrieval pipeline,
\texttt{ChunkService} and the peer request logic, is identical on all platforms.

\par The \texttt{pubspec.yaml} file reflects this split by including both packages
and their platform-specific native library dependencies:

\begin{itemize}
  \item \texttt{just\_audio} with \texttt{just\_audio\_platform\_interface} for
        Android, iOS, and web.
  \item \texttt{media\_kit} with \texttt{media\_kit\_libs\_linux},
        \texttt{media\_kit\_libs\_macos\_audio}, and
        \texttt{media\_kit\_libs\_windows\_audio} for the respective desktop
        platforms.
\end{itemize}

\par Abstracting the player behind the existing \texttt{AppAudioService} interface
meant that no change was needed in the UI layer or in the chunk retrieval layer
when switching between the two backends.  The \texttt{AudioPlayer} instance is
injected into \texttt{AppAudioService} at construction time, so the
platform-specific app class simply passes the correct player implementation to the
service factory.  The UI only interacts with \texttt{AppAudioService} methods such
as \texttt{play()}, \texttt{pause()}, and \texttt{seek()}, and is completely
unaware of which backend is running below.

\subsection{Limitations on the Web Platform}
\label{subsec:just-audio-web}

\par The web platform introduces a fundamentally different problem.  Web browsers
run Flutter applications as JavaScript inside a sandboxed environment.  The
browser's audio APIs, specifically the \texttt{<audio>} element and the Web Audio
API, expect to receive audio from a URL.  The browser makes HTTP range requests to
that URL to load segments of the file progressively, which is the standard way
audio streaming works on the web.

\par This means the \texttt{StreamAudioSource} abstraction used on native platforms
does not apply on the web: the browser directly requests byte ranges from a URL,
and there is no hook at the Dart level to intercept those requests and serve bytes
from the chunk cache instead.

\par One possible solution would be to create an object URL from a full in-memory
\texttt{Blob} of the audio file, effectively downloading the entire song into RAM
before playback starts.  This is unacceptable for a streaming system where the
whole point is to start playback immediately from a partial download.

\par Another possible solution would be to start a local HTTP server inside the
browser tab.  However, browsers do not allow web pages to listen on ports, so this
is not feasible in the browser sandbox.

\par The solution chosen is to use a \emph{Service Worker}, a JavaScript file that
the browser registers as a network proxy for the origin.  The Service Worker
intercepts all HTTP requests that match a specific URL pattern and can respond to
them with any data it chooses, including data assembled from the chunk cache.
Figure~\ref{fig:sw-flow} shows how the Service Worker sits between the browser's
audio element and the Flutter application.

\begin{figure}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=1.0cm and 2.6cm,
  box/.style={rectangle, draw, rounded corners=4pt,
              minimum width=3.2cm, minimum height=0.85cm,
              font=\small, align=center},
  sw/.style={box, fill=red!12},
  flutter/.style={box, fill=blue!10},
  browser/.style={box, fill=orange!12},
  arr/.style={-{Stealth}, thick}
]
  \node[browser]  (audio) {Browser\\Audio Element};
  \node[sw, right=of audio] (sw) {Service Worker\\(p2p-worker.js)};
  \node[flutter, right=of sw] (bridge) {WebP2PBridge\\(Dart)};
  \node[flutter, right=of bridge] (chunk) {ChunkService\\(Dart)};

  \draw[arr] (audio) -- node[above, font=\scriptsize]{HTTP range req.} (sw);
  \draw[arr] (sw) -- node[above, font=\scriptsize]{postMessage} (bridge);
  \draw[arr] (bridge) -- node[above, font=\scriptsize]{getChunk(i)} (chunk);
  \draw[arr] (chunk.south) to[bend left=30]
             node[below, font=\scriptsize]{chunk bytes} (bridge.south);
  \draw[arr] (bridge.south) to[bend left=30]
             node[below, font=\scriptsize]{postMessage} (sw.south);
  \draw[arr] (sw.west) to[bend left=35]
             node[above, font=\scriptsize, xshift=-0.2cm]{HTTP 206 response} (audio.east);
\end{tikzpicture}%
}
\caption{Web audio delivery path: the Service Worker intercepts browser HTTP range
         requests and routes them through \texttt{WebP2PBridge} and
         \texttt{ChunkService}.}
\label{fig:sw-flow}
\end{figure}

\par The Service Worker file, \texttt{p2p-worker.js}, is registered at browser
startup.  It listens for \texttt{fetch} events on URLs that match the pattern
\texttt{/music-player/p2p-stream/\{fileHash\}}.  When a matching request arrives,
the Service Worker sends a \texttt{P2P\_CHUNK\_REQUEST} message to the Flutter
application's main thread via the \texttt{postMessage} API, containing the
\texttt{reqId}, the \texttt{fileHash}, and the HTTP \texttt{Range} header string.
It then suspends the HTTP response promise and waits.

\par The Flutter application listens for this message through the
\texttt{WebP2PBridge} class.  When a \texttt{P2P\_CHUNK\_REQUEST} arrives, the
bridge parses the range string, converts the byte range to chunk indices, calls
\texttt{ChunkService.getChunk} for each required chunk, assembles the bytes into a
contiguous buffer, and posts a \texttt{P2P\_CHUNK\_RESPONSE} message back to the
Service Worker with the assembled bytes, the byte range metadata, and a flag
indicating whether any of the chunks came from a peer.  The Service Worker resolves
the suspended HTTP promise with a \texttt{206 Partial Content} response containing
the correct \texttt{Content-Range} header, which the browser's audio element
consumes as a normal progressive range response.

\par This approach means the audio library on the web is just the browser's native
audio element; no special Flutter audio library is needed.  The chunk retrieval
and peer-assisted delivery logic is identical to the native path: the same
\texttt{ChunkService} and \texttt{WebRTCService} classes are used, the same
integrity verification runs, and the same peer ranking applies.  Only the delivery
mechanism at the boundary with the audio element is different.

\par One practical detail worth noting is that the Service Worker caps each
response at 512\,000 bytes even if the browser requests a larger range.  This is
because some browsers, particularly those based on Chromium, send very large initial
range requests (sometimes requesting the entire file) when first loading an audio
source.  Returning a 512\,000-byte response causes the browser to continue with
smaller, more granular range requests, which map cleanly onto the chunk boundaries
and allow the P2P pipeline to serve them efficiently.

\par One limitation of this approach is that the Service Worker requires an HTTPS
connection in production because browsers refuse to register Service Workers on
insecure origins, with the exception of \texttt{localhost} for development.  This
is not a practical constraint for a deployed streaming service, which would always
use HTTPS.

\par A second limitation is that because the web platform uses an in-memory chunk
cache, cached chunks are lost when the browser tab is closed.  This limits the
contribution of the web client as a P2P seeder between sessions compared to the
desktop client, which persists its chunk cache to disk via the
\texttt{IOChunkCacheRepository}.

% ─────────────────────────────────────────────────────────────────────────────
\section{Client-Side Chunk Retrieval}
\label{sec:chunk-service}

\par On the client the \texttt{ChunkService} class encapsulates the full logic for
obtaining a single chunk, regardless of whether that chunk originates from the
server, a peer, or a local cache.  Each \texttt{ChunkService} instance is bound to
one song and is created by \texttt{AppAudioService} the first time a song is queued
for playback.

\subsection{Caching Architecture}
\label{subsec:caching}

\par Three caching layers are stacked in order of access speed:

\begin{enumerate}
  \item \textbf{Hot RAM cache}: a map of at most 15 entries keyed by chunk index.
        Entries are evicted using a first-in first-out policy when the map exceeds
        capacity.  This layer absorbs repeated reads for the same chunk that occur
        when the audio player issues overlapping range requests.  The limit of 15
        entries was chosen empirically: it is large enough to hold the
        forward-looking prefetch window for a typical seek, but small enough to
        avoid excessive memory pressure on mobile devices where RAM is a
        constrained resource.

  \item \textbf{Disk cache}: a \texttt{ChunkCacheRepository} that stores chunks as
        binary files under a per-song directory.  Writes are asynchronous; the
        chunk is placed in the RAM cache first and persisted in the background so
        that the caller receives the data immediately without waiting for disk I/O.
        On the web platform, the disk cache is replaced by the in-memory repository
        described in Section~\ref{subsec:repo-split}.

  \item \textbf{Active request deduplication}: a map from chunk index to a
        \texttt{Future<Uint8List>}.  If two callers concurrently request the same
        chunk, the second caller attaches to the existing future rather than issuing
        a duplicate network or peer request.  This is particularly important during
        seeking, when the audio player may issue several overlapping range requests
        that all require the same chunk.
\end{enumerate}

\subsection{Prefix Threshold Calculation}
\label{subsec:prefix-calc}

\par The prefix threshold, the number of chunks that are always fetched from the
server, is calculated dynamically rather than hardcoded.  The formula is:

\begin{equation}
  N_{\text{prefix}} = \max\!\bigl(1,\ \text{round}(N_{\text{total}} \times 0.05)\bigr),
  \label{eq:prefix-calc}
\end{equation}

where $N_{\text{total}}$ is the total number of chunks in the song.  The prefix is
always at least 5\% of the song.  For a typical 4-minute song at 320~kbps,
$N_{\text{total}} \approx 120$ and the prefix is 6 chunks, which is approximately
384~KiB or about 9.4 seconds of audio.  For a shorter 1-minute track,
$N_{\text{total}} \approx 30$ and the minimum of 1 is applied.

\par The dynamic threshold is preferable to a fixed value because the goal is to
guarantee that enough data is available from the server during the WebRTC
negotiation phase.  A fixed count of, say, 8 chunks would represent a much larger
fraction of a very short track and would reduce the peer offload ratio for short
content.  The 5\% formula keeps the server cost proportional to song length.

\subsection{Fetch Decision Logic}
\label{subsec:fetch-logic}

\par The core method \texttt{\_fetchChunkLogic(int index)} implements a
three-branch decision tree, illustrated in Figure~\ref{fig:chunk-fetch-flow}:

\begin{enumerate}
  \item If the chunk index falls within the server prefix range (that is,
        $\texttt{index} < N_{\text{prefix}}$), the chunk is always fetched from
        the origin server.  This guarantees that playback can begin while peer
        connections are still being negotiated.

  \item Otherwise, if at least one peer with the requested chunk index is known,
        a P2P request is issued through \texttt{WebRTCService} with a timeout of
        1700 milliseconds.  If the peer responds in time, the data is
        integrity-verified against the SHA-256 hash from the manifest.  A
        verification failure causes an immediate fallback to the server, and if the
        server copy also fails verification, an exception is thrown to halt
        playback with an explicit integrity error.

  \item If no suitable peer exists, or if the peer request times out, the chunk is
        fetched from the origin server.
\end{enumerate}

\par Every successfully fetched chunk is written to both the disk cache and the RAM
cache, and the peer registry is updated via a \texttt{REGISTER\_CACHE} WebSocket
message so that other peers can subsequently discover this device as a source.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  node distance=0.85cm and 0.1cm,
  startstop/.style={ellipse, draw, fill=gray!20,
                    minimum width=2.4cm, minimum height=0.7cm, font=\small},
  decision/.style={diamond, draw, fill=yellow!20, aspect=2.2,
                   minimum width=3.2cm, font=\small, align=center},
  process/.style={rectangle, draw, fill=blue!10, rounded corners=3pt,
                  minimum width=3.6cm, minimum height=0.75cm,
                  font=\small, align=center},
  good/.style={rectangle, draw, fill=green!15, rounded corners=3pt,
               minimum width=3.6cm, minimum height=0.75cm,
               font=\small, align=center},
  bad/.style={rectangle, draw, fill=red!15, rounded corners=3pt,
              minimum width=3.6cm, minimum height=0.75cm,
              font=\small, align=center},
  arr/.style={-{Stealth}}
]
  \node[startstop] (start) {getChunk(index)};

  \node[decision, below=of start] (inCache) {In RAM\\cache?};
  \node[good, right=2.8cm of inCache] (returnCache) {Return cached data};

  \node[decision, below=of inCache] (isPrefix) {index $<$ prefix?};
  \node[process, right=2.8cm of isPrefix] (fetchServer1) {Fetch from\\server};

  \node[decision, below=of isPrefix] (hasPeers) {Peers with\\this chunk?};
  \node[process, right=2.8cm of hasPeers] (fetchServer2) {Fetch from\\server};

  \node[process, below=of hasPeers] (p2pReq)  {Request chunk\\via WebRTC};
  \node[decision, below=of p2pReq]  (timeout) {Response\\within 1.7\,s?};
  \node[process, right=2.8cm of timeout] (fetchServer3) {Fetch from\\server};

  \node[decision, below=of timeout] (verify) {Hash OK?};
  \node[bad, right=2.8cm of verify] (integrity) {Integrity\\error};

  \node[good, below=of verify] (cache) {Cache \&\\return data};

  % edges
  \draw[arr] (start)      -- (inCache);
  \draw[arr] (inCache)    -- node[left, font=\scriptsize]{No}  (isPrefix);
  \draw[arr] (inCache)    -- node[above, font=\scriptsize]{Yes} (returnCache);
  \draw[arr] (isPrefix)   -- node[left, font=\scriptsize]{No}  (hasPeers);
  \draw[arr] (isPrefix)   -- node[above, font=\scriptsize]{Yes} (fetchServer1);
  \draw[arr] (fetchServer1.south) |- (verify.east);
  \draw[arr] (hasPeers)   -- node[left, font=\scriptsize]{No}  (p2pReq);
  \draw[arr] (hasPeers)   -- node[above, font=\scriptsize]{No peers} (fetchServer2);
  \draw[arr] (fetchServer2.south) |- (verify.east);
  \draw[arr] (p2pReq)     -- (timeout);
  \draw[arr] (timeout)    -- node[left, font=\scriptsize]{Yes} (verify);
  \draw[arr] (timeout)    -- node[above, font=\scriptsize]{No}  (fetchServer3);
  \draw[arr] (fetchServer3.south) |- (verify.east);
  \draw[arr] (verify)     -- node[left, font=\scriptsize]{OK}  (cache);
  \draw[arr] (verify)     -- node[above, font=\scriptsize]{Fail} (integrity);
\end{tikzpicture}
\caption{Chunk fetch decision flow inside \texttt{ChunkService.\_fetchChunkLogic}.}
\label{fig:chunk-fetch-flow}
\end{figure}

\subsection{Staggered Peer Fan-Out}
\label{subsec:fanout}

\par When peers are available, the \texttt{\_requestFromPeers} method does not
simply ask one peer and wait.  It ranks the available peers by their score (see
Section~\ref{subsec:peer-ranking}) and picks the top three.  A request is sent to
the best peer immediately.  If the first response has not arrived within
150 milliseconds, a request is also sent to the second peer.  After another
150 milliseconds, if the chunk is still not resolved, a request is sent to the
third peer.  The first peer to respond completes the shared
\texttt{Completer<Uint8List>} and the other responses are ignored because the
completer is already resolved.

\par This staggered fan-out strategy reduces the latency of the common case (the
best peer responds quickly) while still providing a fast recovery path when the
best peer is slow or temporarily unavailable.  The 150-millisecond step was chosen
as a compromise: short enough to trigger the backup quickly if the first peer is
unavailable, but long enough to avoid generating unnecessary traffic when the first
peer is simply a little slow.

\subsection{Look-Ahead Prefetching}
\label{subsec:prefetch}

\par To prevent starvation of the audio decoder, the \texttt{ChunkService}
maintains a \texttt{\_preloadAhead} constant of 16.  After each call to
\texttt{getChunk(i)}, the service asynchronously initiates fetches for chunks
$i+1$ through $i+15$, subject to the deduplication map preventing re-issuance.
These prefetch operations run concurrently so the peer timeout limits only the
latency of a single chunk fetch, not the aggregate prefetch window.

\par The lookahead window of 16 chunks represents approximately 1~MiB of audio
data, which corresponds to about 25 seconds at 320~kbps.  This is large enough
that even a worst-case scenario of one peer timeout per chunk for all 16 prefetch
chunks would still not exhaust the buffer before the last prefetch resolves via
server fallback.  In practice, the buffer grows faster than it drains when most
chunks come from peers that respond well within the timeout.

% ─────────────────────────────────────────────────────────────────────────────
\section{P2P Audio Source and Range Request Handling}
\label{sec:audio-source}

\par The Flutter \texttt{just\_audio} library consumes audio through an abstract
\texttt{AudioSource} interface.  For P2P streaming the application provides
\texttt{P2PChunkedAudioSource}, which extends \texttt{StreamAudioSource} and
translates HTTP-style range requests into calls to \texttt{ChunkService.getChunk}.

\par When the audio player issues a range request for bytes $[s, e)$, the source:

\begin{enumerate}
  \item Clamps the range to $[0, \texttt{totalBytes})$.
  \item Computes the first and last chunk indices:
        $i_{\min} = \lfloor s / C \rfloor$,
        $i_{\max} = \lfloor (e-1) / C \rfloor$,
        where $C = 65\,536$ bytes is the chunk size.
  \item Iterates from $i_{\min}$ to $i_{\max}$, awaiting each chunk future in
        turn while simultaneously scheduling a speculative fetch for chunk $i + 1$
        to overlap network latency with audio decoding.
  \item For each chunk, computes the byte-level slice:
        $\texttt{sliceStart} = \max(0,\ s - i \cdot C)$,
        $\texttt{sliceEnd} = \min(|\texttt{chunk}|,\ e - i \cdot C)$,
        and yields the sub-list to the output stream.
\end{enumerate}

\par The response object declares \texttt{rangeRequestsSupported: true} and sets
the \texttt{Content-Length} header to $e - s$, allowing the audio player to issue
fine-grained seeks without re-downloading the entire file.  Setting this flag is
important because without it some audio players assume the source is non-seekable
and disable the seek bar in the UI.

% ─────────────────────────────────────────────────────────────────────────────
\section{WebRTC Peer Management}
\label{sec:webrtc}

\par The \texttt{WebRTCService} on the client side manages the lifecycle of all
peer-to-peer connections.  It is constructed with references to the signaling
WebSocket and two callbacks: \texttt{onChunkReceived} (invoked when binary chunk
data arrives over a data channel) and \texttt{onChunkRequested} (invoked when a
peer requests a chunk that this device may be able to serve).

\subsection{Connection Establishment}
\label{subsec:conn-establish}

\par Peer connections are established using the standard WebRTC offer/answer
handshake.  When a \texttt{PEER\_BUFFER\_MAP} response arrives from the signaling
server, the service iterates the map and calls
\texttt{\_createPeerConnection(peerId, initiator: true)} for each newly discovered
peer.  This method:

\begin{enumerate}
  \item Creates an \texttt{RTCPeerConnection} with the configured STUN servers
        (\texttt{stun.l.google.com:19302} and \texttt{stun1.l.google.com:19302}).
        Two servers are included so that if one is temporarily unreachable the ICE
        negotiation can still complete.

  \item Creates an ordered, reliable \texttt{RTCDataChannel} labelled
        \texttt{"chunk\_transfer"}.  The ordered and reliable configuration mirrors
        TCP semantics, which is important because audio chunks must arrive complete
        and uncorrupted.  A partially delivered 64~KiB block cannot be verified by
        the SHA-256 check and would be discarded, wasting both the transfer cost
        and the peer timeout budget.

  \item Generates an SDP offer, sets it as the local description, and transmits it
        to the signaling server as an \texttt{OFFER} message, including an
        \texttt{offerId} token that the remote peer echoes back in its answer.
        This token is used to detect stale answers that arrive after a connection
        has been re-established, preventing old answers from overwriting the state
        of a fresh connection.
\end{enumerate}

\par The remote peer receives the offer, creates a mirrored connection with
\texttt{initiator: false}, sets the remote description, and replies with an
\texttt{ANSWER} message.  Both sides then exchange \texttt{ICE\_CANDIDATE}
messages until connectivity is established.  ICE candidates that arrive before the
remote description is set are queued in \texttt{\_iceQueues} and applied once the
description is available, because the WebRTC API requires the remote description
to be present before candidates can be added.

\par A known edge case in WebRTC negotiation is \emph{glare}: both sides send an
offer simultaneously, before either has had a chance to reply with an answer.  This
can happen when two clients discover each other at nearly the same moment, for
example when both join a room listening to the same song.  The implementation
resolves glare deterministically: the peer with the lexicographically larger device
identifier keeps its offer and ignores the incoming one; the other peer discards
its local offer and processes the incoming one as the responder.  This resolves the
ambiguity without requiring a server-side arbitration step.

\subsection{Chunk Exchange Protocol}
\label{subsec:chunk-protocol}

\par Once the data channel is open, chunk requests and responses follow a custom
binary protocol.  A \emph{request} is sent as a UTF-8 text message with the
format:

\begin{center}
  \texttt{REQUEST\_CHUNK:\{fileHash\}:\{chunkIndex\}}
\end{center}

\par A \emph{response} is a binary message.  If the total chunk data fits within
the safe DataChannel message limit of 16~KiB, a single packet is sent.  The header
is 68 bytes: the first 64 bytes are the UTF-8 encoding of the file hash (which is
always a 64-character hexadecimal string), and bytes 64--67 are the chunk index
encoded as a big-endian 32-bit unsigned integer.  The remaining bytes are the raw
audio data.

\par If the chunk data exceeds the 16~KiB limit, it is split into fragments.  Each
fragment uses a 77-byte header: a 4-byte magic sequence \texttt{P2PF} to
distinguish fragments from single-packet messages, a 1-byte version number, the
64-byte file hash, a 4-byte chunk index, a 2-byte fragment index, and a 2-byte
total fragment count.  The receiving side buffers incoming fragments in a
\texttt{\_ChunkAssembly} structure until all fragments for a given
\texttt{(peerId, fileHash, chunkIndex)} tuple have arrived, then reassembles and
delivers the complete chunk.

\par This fragmentation protocol was added because some implementations of the
WebRTC DataChannel API, particularly in certain versions of the underlying native
library on Android, refuse to send messages larger than around 16~KiB.  Without
fragmentation, any chunk that happened to be produced by an encoder at a slightly
higher bitrate, resulting in a chunk slightly over 16~KiB, would fail silently on
those platforms.  The magic bytes allow the receiving side to distinguish
fragmented packets from legacy single-packet responses without breaking backward
compatibility.

\par If a chunk request arrives while the data channel is still in the
\texttt{connecting} state, the request is enqueued in
\texttt{\_pendingChunkRequests} and flushed via \texttt{\_drainPendingRequests} as
soon as the channel transitions to \texttt{open}.  This prevents race conditions
during the brief window between peer discovery and channel readiness.

\subsection{Peer Ranking and Scoring}
\label{subsec:peer-ranking}

\par Not all peers are equal.  A peer on the same local network will typically
deliver a chunk in under 10 milliseconds, while a peer on the other side of the
world may take several hundred milliseconds.  To exploit the fastest available
peers, each remote peer is tracked by a \texttt{\_PeerStats} object that maintains
two exponentially weighted moving averages:

\begin{itemize}
  \item \textbf{Round-trip time (RTT)}: updated every 30 seconds via a
        \texttt{DC\_PING}/\texttt{DC\_PONG} exchange over the data channel.  The
        keepalive timer sends a \texttt{DC\_PING:\{timestamp\}} text message; the
        remote peer echoes it back as \texttt{DC\_PONG:\{timestamp\}}, and the RTT
        is computed from the difference.  The keepalive also prevents idle
        DataChannel connections from being closed by intermediate NAT devices that
        drop inactive UDP associations.

  \item \textbf{Throughput}: updated whenever a chunk is delivered, as the ratio
        of the chunk size to the elapsed time between the request being sent and
        the response being received.
\end{itemize}

\par The peer score is:

\begin{equation}
  \text{score} = \frac{\text{successRate} \times \text{throughput} \times 1000}
                      {\text{RTT} + 1},
  \label{eq:peer-score}
\end{equation}

where $\text{successRate}$ is the fraction of requests to this peer that were
answered successfully (not timed out), \texttt{throughput} is the exponentially
averaged bytes per millisecond, and RTT is the exponentially averaged round-trip
time in milliseconds.  The $+1$ in the denominator prevents division by zero for
newly seen peers.

\par The exponential smoothing factor $\alpha = 0.2$ gives recent measurements 20\%
weight and the historical average 80\% weight.  This makes the score stable enough
to avoid frequent re-ranking due to transient fluctuations, while still reacting
to sustained changes in peer performance within a few measurement intervals.

\par When \texttt{ChunkService} calls \texttt{getSortedPeersForSong(fileHash)}, the
service returns the list of peers sorted by score in descending order, filtered to
those that have registered chunks for the requested file.  The top three are used
in the staggered fan-out described in Section~\ref{subsec:fanout}.

\subsection{Upload Limits and Network Mode}
\label{subsec:upload-limits}

\par On mobile platforms, acting as a peer seeder consumes the user's cellular or
Wi-Fi data allowance.  To give users control over this, \texttt{WebRTCService}
checks two settings before sending a chunk to another peer:

\begin{enumerate}
  \item \textbf{Peer network mode}: the user can configure the application to
        participate in P2P only on Wi-Fi (\texttt{mode = 0}), only on cellular
        (\texttt{mode = 1}), or on both (\texttt{mode = 2}).  The
        \texttt{connectivity\_plus} package is used to check the current connection
        type.  If the current connection does not match the configured mode, the
        chunk send is silently skipped and the requesting peer falls back to the
        server.

  \item \textbf{Monthly upload cap}: separately configurable for Wi-Fi and cellular.
        The service accumulates the total bytes sent per calendar month and refuses
        to send further chunks once the cap is reached.  The counters reset at the
        start of each calendar month.  The month boundary is stored as a
        \texttt{year * 100 + month} integer so that it survives application
        restarts.
\end{enumerate}

\par On desktop platforms and on the web these checks are bypassed because desktop
machines are typically on unmetered connections, and because the web platform
cannot reliably distinguish cellular from Wi-Fi traffic.

% ─────────────────────────────────────────────────────────────────────────────
\section{Audio Service and Playback State Management}
\label{sec:audio-service}

\par \texttt{AppAudioService} is the top-level coordinator for audio playback on
the client.  It owns the \texttt{AudioPlayer} instance (either \texttt{just\_audio}
or the \texttt{media\_kit} equivalent depending on the platform), manages the song
queue, and bridges playback events to the data synchronisation and peer registry
subsystems.

\subsection{Queue and Audio Source Selection}
\label{subsec:queue}

\par When the user starts a song, \texttt{setQueueAndPlay} atomically replaces the
queue and calls \texttt{\_buildAudioSource(song)}.  The factory branches on three
cases:

\begin{itemize}
  \item If the song has a local file path (for example, a file the user imported
        from disk on desktop), a \texttt{Uri.file} source is returned and no
        network I/O is needed at all.

  \item If the platform is web, the audio source is a URL under the
        \texttt{/music-player/p2p-stream/} scheme that the Service Worker
        intercepts.  The \texttt{AppAudioService} notifies the \texttt{WebP2PBridge}
        of the current song via a callback so that the bridge can associate incoming
        Service Worker requests with the correct \texttt{ChunkService} instance.

  \item Otherwise (native streaming), a \texttt{P2PChunkedAudioSource} is
        constructed, which triggers a \texttt{ChunkService} instantiation and a
        manifest fetch on the first range request.
\end{itemize}

\par Both queues (the normal queue and the shuffled queue) are kept in sync at all
times.  When shuffle is toggled, the shuffled queue is rebuilt from the normal
queue with the current song placed at the front of the shuffle, so that the user
does not jump away from the currently playing song unexpectedly.

\subsection{Proactive Prefix Caching}
\label{subsec:proactive-cache}

\par To reduce the latency of switching to the next song in the queue, the service
runs \texttt{\_proactivelyCachePrefixes} each time the 30-second timer fires while
a song is playing.  This method inspects the next five songs in the queue and, for
each one, issues a server prefix fetch.  Because the prefix is a single HTTP call
returning less than 400~KiB on average, it can be stored in the disk cache well
before the user reaches that position in the queue.  This makes the subsequent
peer-negotiation phase overlap with warm cache hits rather than cold fetches,
effectively hiding the WebRTC connection setup latency from the user.

\par The proactive cache covers only five songs ahead rather than the entire queue
for bandwidth reasons.  Five songs correspond to approximately 2~MiB of background
data.  For longer queues the assumption is that the user will not advance through
the queue so quickly that the prefetching cannot keep up; the 30-second timer
provides ample time to warm the cache for upcoming songs.

\subsection{Stuck-Playback Detection}
\label{subsec:stuck}

\par Network jitter or slow peer responses can occasionally stall the audio decoder
without triggering a formal error state in the player library.
\texttt{AppAudioService} runs a watchdog timer that samples the playback position
every 500 milliseconds.  If the position has not advanced after two consecutive
samples (a 1000-millisecond window) while the player is nominally in the
\texttt{playing} state, the service reloads the audio source from the current
position.  This forces the player to re-issue range requests, which in turn
triggers fresh chunk fetches that may select a different peer or fall through to
the server, resolving the stall.

\par The watchdog is started each time \texttt{play()} is called and exits as soon
as the position advances or the player leaves the \texttt{playing} state.  It does
not run on a periodic timer to avoid accumulating concurrent watchdog coroutines in
long sessions; each call to \texttt{play()} spawns exactly one watchdog that runs
to completion.

\par One reason this detection is necessary is that \texttt{just\_audio} and
\texttt{media\_kit} have different behaviour when the byte stream from the source
stops producing data.  On some platforms the player silently waits indefinitely;
on others it emits a \texttt{buffering} state.  The watchdog provides a
platform-independent safety net that does not depend on the player emitting a
specific state transition.

\subsection{Persistent Playback State}
\label{subsec:playback-state}

\par Every 30 seconds while playback is active, \texttt{pushStateToServer}
serialises the current queue as a JSON array of song IDs, the playback position in
milliseconds, and the shuffle and repeat flags into a \texttt{PlaybackStateDto},
which is sent to the \texttt{PlaybackRestService} REST endpoint.  On application
startup, \texttt{restoreFromServerState} fetches the last saved state and
reconstructs the queue and position, enabling seamless handoff between devices and
resumption after application restarts.

\par The 30-second interval is a deliberate trade-off.  Saving more frequently
would reduce the maximum data loss in the event of a crash, but would also generate
more write traffic on the server.  At 30 seconds, the worst-case position loss is
30 seconds of playback, which is acceptable for a music application where the user
can skip back a few seconds manually.

% ─────────────────────────────────────────────────────────────────────────────
\section{End-to-End Flows}
\label{sec:flows}

\par This section consolidates the subsystem descriptions above into three
end-to-end flow diagrams that emphasise the temporal ordering of operations and
the interactions between components.

\subsection{Playback Initialisation Flow}
\label{subsec:flow-init}

\par Figure~\ref{fig:flow-init} shows the sequence of operations from the moment a
user selects a song to the moment audio begins playing.  The critical property is
that playback starts after the first byte range is served from server prefix chunks
while peer connection establishment continues asynchronously in the background.
The user experiences no delay related to WebRTC negotiation.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  node distance=0.7cm,
  box/.style={rectangle, draw, fill=blue!8, rounded corners=3pt,
              minimum width=10cm, minimum height=0.7cm,
              font=\small, align=flush left, text width=9.6cm,
              inner sep=4pt},
  phase/.style={rectangle, draw=none, fill=gray!15, rounded corners=2pt,
                minimum width=10cm, minimum height=0.55cm,
                font=\small\bfseries, align=center},
  arr/.style={-{Stealth}, thick, gray!60}
]
  \node[phase]                  (p1) {Phase 1: Manifest and Prefix};
  \node[box, below=0.1cm of p1] (s1) {1. User selects song $\rightarrow$ \texttt{AppAudioService.setQueueAndPlay}};
  \node[box, below=0.2cm of s1] (s2) {2. Audio source built: native $\rightarrow$ \texttt{P2PChunkedAudioSource}; web $\rightarrow$ stream URL};
  \node[box, below=0.2cm of s2] (s3) {3. \texttt{ChunkService.loadManifest()} $\rightarrow$ HTTP GET /manifest};
  \node[box, below=0.2cm of s3] (s4) {4. Server returns hash list, chunk count, total bytes};
  \node[box, below=0.2cm of s4] (s5) {5. Client begins playback; server serves prefix chunks on first range request};

  \node[phase, below=0.4cm of s5] (p2) {Phase 2: Peer Discovery (concurrent with Phase 1)};
  \node[box, below=0.1cm of p2]  (s6) {6. \texttt{ChunkService} sends \texttt{DISCOVER\_PEERS} over WebSocket};
  \node[box, below=0.2cm of s6]  (s7) {7. Server queries Redis, returns \texttt{PEER\_BUFFER\_MAP}};
  \node[box, below=0.2cm of s7]  (s8) {8. \texttt{WebRTCService} initiates \texttt{RTCPeerConnection} per peer};
  \node[box, below=0.2cm of s8]  (s9) {9. OFFER/ANSWER/ICE exchanged via signaling server};

  \node[phase, below=0.4cm of s9]  (p3) {Phase 3: Peer-Assisted Streaming};
  \node[box, below=0.1cm of p3]   (s10) {10. Audio player issues range request $\rightarrow$ source calls \texttt{getChunk(i)}};
  \node[box, below=0.2cm of s10]  (s11) {11. Prefix chunks: served from server; cached and registered};
  \node[box, below=0.2cm of s11]  (s12) {12. Non-prefix chunks: P2P first (1.7 s timeout), server fallback};
  \node[box, below=0.2cm of s12]  (s13) {13. Each chunk verified against manifest hash, cached, registered};

  \draw[arr] (s1.south) -- (s2.north);
  \draw[arr] (s2.south) -- (s3.north);
  \draw[arr] (s3.south) -- (s4.north);
  \draw[arr] (s4.south) -- (s5.north);
  \draw[arr] (s6.south) -- (s7.north);
  \draw[arr] (s7.south) -- (s8.north);
  \draw[arr] (s8.south) -- (s9.north);
  \draw[arr] (s10.south) -- (s11.north);
  \draw[arr] (s11.south) -- (s12.north);
  \draw[arr] (s12.south) -- (s13.north);
\end{tikzpicture}
\caption{Playback initialisation flow from song selection to peer-assisted streaming.}
\label{fig:flow-init}
\end{figure}

\subsection{Web Playback Flow}
\label{subsec:flow-web}

\par On the web platform the path between the audio element and the
\texttt{ChunkService} involves an extra hop through the Service Worker.
Figure~\ref{fig:flow-web} shows the message sequence for a single byte-range
request on the web.

\begin{figure}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=1.0cm and 2.2cm,
  box/.style={rectangle, draw, rounded corners=3pt,
              minimum width=2.4cm, minimum height=0.8cm,
              font=\small, align=center},
  br/.style={box, fill=orange!15},
  sw/.style={box, fill=red!12},
  dart/.style={box, fill=blue!10},
  arr/.style={-{Stealth}, thick},
  darr/.style={-{Stealth}, thick, dashed}
]
  \node[br]  (audio) {Browser\\Audio Element};
  \node[sw,  right=2cm of audio]  (swNode) {Service Worker\\(p2p-worker.js)};
  \node[dart, right=2cm of swNode] (bridge) {WebP2PBridge};
  \node[dart, right=2cm of bridge] (chunk)  {ChunkService};

  % Request path
  \draw[arr] (audio)  -- node[above, font=\scriptsize]{Range: bytes=X-Y}  (swNode);
  \draw[darr] (swNode) -- node[above, font=\scriptsize]{P2P\_CHUNK\_REQUEST} (bridge);
  \draw[arr] (bridge) -- node[above, font=\scriptsize]{getChunk(i)} (chunk);

  % Response path
  \draw[darr] (chunk.south)  to[bend left=30]
              node[below, font=\scriptsize]{Uint8List} (bridge.south);
  \draw[darr] (bridge.south) to[bend left=30]
              node[below, font=\scriptsize]{P2P\_CHUNK\_RESPONSE} (swNode.south);
  \draw[arr]  (swNode.west)  to[bend left=35]
              node[above, font=\scriptsize, xshift=0.5cm]{HTTP 206 Partial} (audio.east);
\end{tikzpicture}%
}
\caption{Web playback request flow: the Service Worker proxies browser HTTP range
         requests to the Flutter \texttt{WebP2PBridge}, which assembles bytes from
         \texttt{ChunkService}.}
\label{fig:flow-web}
\end{figure}

\subsection{Chunk Delivery Decision}
\label{subsec:flow-chunk}

\par The chunk delivery path shown in Figure~\ref{fig:chunk-fetch-flow} is the
innermost hot path of the system.  Its design reflects two competing constraints:
\emph{latency} (the audio decoder cannot buffer ahead if chunk fetches are slow)
and \emph{server cost} (the goal is to minimise origin-server bandwidth).

\par The dynamic prefix threshold resolves this tension by guaranteeing a baseline
supply of data from a reliable source during the WebRTC connection setup window.
As long as the prefix covers more audio time than the ICE negotiation takes, the
transition from server-delivered prefix to peer-delivered suffix is seamless.

\par The 1700-millisecond P2P timeout was chosen empirically from testing in
several network environments.  The value in Chapter~\ref{chap:ch2} was derived
theoretically as 1000 milliseconds, but in practice ICE negotiation in cellular
NAT environments took closer to 1200--1500 milliseconds.  The extra margin reduces
the number of unnecessary server fallbacks without meaningfully increasing the risk
of audible stalls, because the 16-chunk lookahead provides a large enough buffer.

\subsection{Library Synchronisation Flow}
\label{subsec:flow-sync}

\par Figure~\ref{fig:flow-sync} shows the multi-device library sync triggered by
a user action such as liking a song or deleting a track from the library.

\begin{figure}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=1.1cm and 3.0cm,
  box/.style={rectangle, draw, rounded corners=3pt,
              minimum width=2.8cm, minimum height=0.8cm,
              font=\small, align=center},
  devA/.style={box, fill=blue!12},
  server/.style={box, fill=orange!15},
  redis/.style={box, fill=red!12},
  devB/.style={box, fill=green!12},
  arr/.style={-{Stealth}, thick}
]
  % Column headers
  \node[devA]    (a1) {Device A\\Action};
  \node[server, right=2.4cm of a1]  (srv) {Server};
  \node[redis, right=2.4cm of srv]  (red) {Redis};
  \node[devB, right=2.4cm of red]   (b1)  {Device B};

  % Steps
  \node[devA, below=of a1]   (a2) {POST /sync\\localChanges};
  \node[server, below=of srv] (sv2) {Apply changes\\Query deltas};
  \node[server, below=of sv2] (sv3) {Publish\\SYNC\_TRIGGER};
  \node[redis, below=1.8cm of red]  (r1)  {Broadcast to\\all instances};
  \node[devB, below=3.0cm of b1]   (b2)  {Receive trigger\\POST /sync};
  \node[devB, below=of b2]   (b3)  {Apply server\\deltas locally};

  % Arrows
  \draw[arr] (a2.east)  -- node[above,font=\scriptsize]{HTTP}   (sv2.west);
  \draw[arr] (sv2.south)-- (sv3.north);
  \draw[arr] (sv3.east) -- node[above,font=\scriptsize]{pub}    (r1.west);
  \draw[arr] (r1.east)  -- node[above,font=\scriptsize]{sub}    (b2.west);
  \draw[arr] (b2.south) -- (b3.north);
  % Response arrow
  \draw[arr] (sv2.west) to[bend right=15]
              node[below,font=\scriptsize]{SyncResponse} (a2.south);
\end{tikzpicture}%
}
\caption{Multi-device library synchronisation triggered by a user action on Device A.}
\label{fig:flow-sync}
\end{figure}

% ─────────────────────────────────────────────────────────────────────────────
\section{Delivery Statistics and Observability}
\label{sec:stats}

\par To validate the effectiveness of the peer-assisted approach, the system
records per-session delivery statistics.  On native platforms, the
\texttt{ChunkService} accumulates counts of chunks served from the local cache,
from peers, and from the origin server.  It fires a \texttt{ChunkDeliveryStats}
value object when the last chunk of a song has been delivered, and a periodic timer
also emits intermediate statistics every 15 seconds so that per-song data is not
lost if the user skips away before the song ends.

\par On the web platform, statistics are collected differently because the
\texttt{ChunkService} does not know which byte ranges the browser audio element is
requesting.  The Service Worker counts how many byte-range responses it marks as
P2P (using the \texttt{isP2P} flag from the \texttt{P2P\_CHUNK\_RESPONSE} message)
and how many come from the server.  It reports these counts back to the Flutter
application at 15-second intervals via a \texttt{P2P\_STATS\_REPORT} message,
which the \texttt{WebP2PBridge} forwards to the statistics reporting service, which
posts them to the REST endpoint.

\par These records allow an operator to observe, for a given deployment, what
fraction of total data volume is served by peers rather than the origin server.
Under favourable conditions, where several listeners are playing the same song at
the same time, the peer offload ratio should approach the theoretical maximum of
$(N - 1) / N$ for a swarm of $N$ simultaneous listeners, limited in practice by
the prefix requirement and connection churn as discussed in
Chapter~\ref{chap:ch2}.
