%\documentclass[12pt]{scrreprt}
\documentclass[12pt]{report} 

% language may be romanian or english (default is english)
% type may be bachelor or master (default is bachelor)
\usepackage[language=english, type=bachelor]{style}

%\geometry{a4paper,top=2.5cm,left=3cm,right=2.5cm,bottom=2.5cm}
%in style
%controlling the appearance of your headers and footers
\usepackage{fancyhdr}
\usepackage{pgfplots}
\usepackage{amsmath} 
\usetikzlibrary{positioning, arrows.meta, shapes.geometric}
\pagestyle{fancy}
\lhead{}
\chead{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.2pt}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\begin{document}

\specialization{COMPUTER SCIENCE}	
\title{Minimizing Cloud Infrastructure Costs in Content Streaming Via Peer-Assisted Delivery and Dynamic Prefix Caching}					   
\author{Darius-George Sala}											
\supervisor{Dr. Lect. Prof. Mircea Ioan-Gabriel}				
				
\maketitle


\newpage
\thispagestyle{empty}
\mbox{}
\newpage
\pagenumbering{roman} 

\cleardoublepage
ABSTRACT
\vspace{0.5cm}	
\hrule
\vspace{0.5cm}	
%\cleardoublepage
This thesis studies how the cloud egress cost of on-demand music streaming can be reduced without changing the listener's experience. The proposed solution is a peer-assisted streaming system in which the server delivers a small initial prefix of each song and remains available as a fallback, while the rest of the audio can be delivered directly between listeners through WebRTC. The original contribution is the design, implementation and evaluation of a single cross-platform prototype that combines dynamic prefix caching, SHA-256 chunk verification, peer ranking, server fallback and multi-replica signaling into one working system.

\par Chapter \ref{intro} presents the problem context and related work. It explains why cloud egress is a relevant cost for music streaming, revisits earlier peer-assisted systems such as Spotify's former P2P architecture, and organises the design space around five problems: startup latency, connectivity, storage, integrity and flow control.

\par Chapter \ref{chap:ch2} turns these trade-offs into a theoretical model. It derives the main design choices of the system, including the 64 KiB content-addressed chunk size, the dynamic 5\% server prefix, the use of WebRTC without TURN relays, and adaptive routing instead of adaptive bitrate.

\par Chapter \ref{chap:ch3} describes the implementation of the prototype, called MP33r. The backend is built with Spring Boot, PostgreSQL and Redis, while the client is written in Flutter and runs on Linux, macOS, Windows, Android, iOS and the web. Audio files are split into SHA-256-addressed chunks, stored with server-side deduplication, verified on the client regardless of source, and exchanged between peers over WebRTC data channels.

\par Chapter \ref{chap:ch4} evaluates the system using a headless measurement client. In the measured warm-swarm scenario, two to five listeners received 88.4\% of their bytes from peers, while the origin share remained flat as the number of listeners grew. The 5\% prefix also hid the median 243 ms WebRTC setup time, keeping click-to-sound latency at the level of a normal server stream.

\par Chapter \ref{chap:ch5} concludes that the listeners of a song can act as part of the delivery infrastructure for that song, as long as the system preserves instant playback, verifies untrusted chunks and keeps a reliable server fallback. The evaluation is limited to a controlled single-machine setup, so the results are not presented as production measurements. Instead, they validate the mechanism and the cost model under repeatable conditions. Future work includes wide-area experiments, warm-cache measurements, larger swarms, stronger mixed-platform connectivity and interoperability with local media players or other streaming services.



\tableofcontents


\cleardoublepage
\markboth{}{}
INTRODUCTION
\vspace{0.5cm}
\hrule
\vspace{0.5cm}

Every time a song is played from a cloud-hosted streaming service, the provider pays for the same bytes in two different ways: first to store and process them, and then again when they leave the data centre. That second cost, usually called egress, grows with the number of listeners, not with how much money those listeners bring in. This means that a song becoming popular is good news and bad news at the same time: more people are listening, but the delivery bill is also rising. Large providers can absorb this through scale, contracts and CDNs, but the basic shape of the problem remains very simple. Double the audience, and the origin has to serve roughly double the data.

\par The interesting part is that the problem already contains a possible delivery network. A listener who has just received a piece of a song has a valid copy of that piece, and another listener starting the same song shortly after may need exactly those bytes. If clients can safely pass chunks between themselves, the origin server no longer has to be the source of every byte. It can become the component that starts playback, seeds the swarm and stays available when peers are missing. The hard part is making this happen without turning the user experience into an experiment. Playback still has to start quickly, the same client has to work on desktop, mobile and the web, and a chunk received from another user has to be checked as strictly as a chunk received from the server. This thesis designs, builds and measures a music streaming system around those constraints, moving the origin away from serving $O(N)$ full copies of the same song and toward a small bootstrap-and-fallback role in the ideal case.

\par The thesis is organised into five chapters. Chapter \ref{intro} sets out the problem context and surveys the related work, breaking peer-assisted streaming into five design domains: latency, connectivity, storage, integrity and flow control. It weighs the candidate solutions for each of them and reviews the prior systems that inform the design, without yet committing to one. Chapter \ref{chap:ch2} turns those trade-offs into a formal cost model, derives the system's key parameters, including the 64 KiB chunk and the dynamic 5\% prefix, and commits to the synthesised architecture that the rest of the thesis implements and evaluates. Chapter \ref{chap:ch3} presents the implementation in three parts: the architecture and its layers, the runtime flows together with the testing approach, and the user interface. Chapter \ref{chap:ch4} evaluates the working prototype by measuring egress offload, startup latency and scaling as the audience grows, then places the results next to the systems from the literature. Chapter \ref{chap:ch5} closes the loop with the conclusions and the directions for future work.

\vspace{1cm}
\par The original contributions of this work are the following:
\begin{itemize}
  \item a single Flutter and Spring Boot code base that runs the same peer-assisted streaming client across six platforms;
  \item a dynamic prefix-caching bootstrap, serving roughly 5\% of each song from the origin, that hides the WebRTC connection setup so playback can start while the peer mesh forms in the background;
  \item a 64 KiB content-addressed (SHA-256) chunking scheme that deduplicates storage on the server and lets the client verify every chunk in the same way, whether it arrived from a peer or from the origin;
  \item a WebRTC data-channel peer transport with round-trip time and throughput-based peer ranking, staggered fan-out and an always-present server fallback, deliberately built without a TURN relay so the cost-saving property of the peer model is preserved;
  \item prefix-only predictive prefetching of upcoming queue entries, which removes the gap between songs at a bounded bandwidth cost;
  \item a multi-replica backend whose cross-replica peer signaling is hidden from the client through Redis publish/subscribe;
  \item an end-to-end evaluation, built on a purpose-made headless measurement client, that quantifies the egress saving, validates the cost model, and uncovered a signaling concurrency bottleneck that was capping multi-peer scaling.
\end{itemize}


\newpage
\pagenumbering{arabic}

\input{chapters/chapter1_introduction}
%\addcontentsline{toc}{chapter}{Introducere}
%\addcontentsline{toc}{chapter}{Introduction}

\input{chapters/chapter2}
\input{chapters/chapter3}
\input{chapters/chapter4}
\input{chapters/chapter5}

%\addcontentsline{toc}{chapter}{Concluzii}
%\addcontentsline{toc}{chapter}{Conclusions}

\bibliography{references}

\end{document}
